---
title: "qFIT OpenSafely analysis"
author: "Nick Kennedy"
date: "06/10/2021"
output:
  html_document:
      toc: true
      theme: cerulean
      df_print: kable
      dev: png
      fig_width: 12
      fig_height: 8
params:
  suppress_if_le: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(rlang)
source("sro_gastro_functions.R")
walk(list.files("firsttable/R", full.names = TRUE), source)
flip_first_table <- function(x) {
  x %>% as.data.frame() %>%
    select(-Level) %>%
    column_to_rownames("Variable") %>%
    t() %>%
    as.data.frame()
}
print_ft <- function(x) {
  x %>% 
    select(-Level) %>% 
    knitr::kable()
}
```

```{r load data}
start_date <- ymd("2020-04-17")
input <- read_csv("../output/input.csv", col_types = cols(.default = col_character())) %>% 
  mutate(
    across(c(matches("^(qfit|qfit_p\\d+|ft.*|crc.*|earliest.*)_date$"), colonoscopy_sus, flexi_sig_sus, colonoscopy_primary_care, flexi_sig_primary_care), lubridate::ymd),
    across(c(ft_referral, ft_clinic, crc_diagnosis, registered_in_main_study_period), ~as.logical(as.integer(.x))),
    across(c(qfit, age, fobt_num), as.numeric),
    qfit_orig = if_else(is.na(qfit_date), NA_real_, qfit),
    qfit = qfit +
      case_when(
        qfit_comp == "<" ~ -1,
        qfit_comp == ">" ~  1,
        qfit_comp == "=" ~  0
      ),
    qfit_pos = qfit >= 10,
    qfit_pos_assume_zero_neg = case_when(
      !is.na(qfit) ~ qfit >= 10,
      qfit_orig == 0 ~ FALSE,
      TRUE ~ NA
    ),
    qfit_three_cat = case_when(
      qfit >= 10 ~ "positive",
      qfit < 10 ~ "negative",
      qfit_orig == 0 ~ "null"
    ),
    across(matches("qfit_p[1-3]_date"), list(present = ~!is.na(.x))), 
    earliest_colonoscopy = pmin(colonoscopy_primary_care, colonoscopy_sus, na.rm = TRUE) %>% 
      {if_else(is.finite(.), ., as.Date(NA))},
    earliest_flexi_sig = pmin(flexi_sig_primary_care, flexi_sig_sus, na.rm = TRUE) %>% 
      {if_else(is.finite(.), ., as.Date(NA))},
    earliest_crc_diagnosis = pmin(earliest_crc_diagnosis_ever_date, earliest_crc_admission_ever_date, na.rm = TRUE),
    any_crc_in_window = !is.na(earliest_crc_diagnosis) & earliest_crc_diagnosis >= start_date
  )
primary_data <- input %>%
  mutate(
    across(
      starts_with("qfit"),
      ~if_else(qfit_date <= ymd("2020-12-31"), .x, .x[length(.x) + 1]) # Replace any qFIT data beyond end of 2020 with null values
    )
  ) %>% 
  filter(
    is.na(earliest_crc_diagnosis) |
      earliest_crc_diagnosis >= start_date,
    registered_in_main_study_period
  )

primary_data_over_50 <- primary_data %>% 
  filter(age >= 50)
cat("Data last updated at ", as.character(file.mtime("../output/input.csv")), "\n", sep = "")
cat("Analysis run at ", as.character(Sys.time()), "\n", sep = "")
```

# Top level summary by region

```{r}
ft_options <- first_table_options(include_denom = TRUE, percent_first = TRUE, include_p = FALSE, suppress_if_le = params$suppress_if_le)
qfit_overall_rows <- quos(
  "qFIT done and included here" = !is.na(qfit_date),
  "Age if qFIT done" = first_table_row(age, data_filter = !is.na(qfit_date)),
  "Age if qFIT neg" = first_table_row(age, data_filter = !qfit_pos),
  "Age if qFIT pos" = first_table_row(age, data_filter = qfit_pos)
)
qfit_rows <- quos(
  !!!qfit_overall_rows,
  "qFIT category" = qfit_three_cat,
  "Age" = age,
  "Colorectal cancer" = crc_diagnosis,
  "Colorectal cancer if qFIT < 10" = first_table_row(crc_diagnosis, data_filter = !qfit_pos),
  "Colorectal cancer if qFIT >= 10" = first_table_row(crc_diagnosis, data_filter = qfit_pos),
  "Colorectal cancer if qFIT null" = first_table_row(crc_diagnosis, data_filter = qfit_three_cat == "null"),
  "Colorectal cancer (inc sec care)" = any_crc_in_window,
  "Colorectal cancer (inc sec care) if qFIT < 10" = first_table_row(any_crc_in_window, data_filter = !qfit_pos),
  "Colorectal cancer (inc sec care) if qFIT >= 10" = first_table_row(any_crc_in_window, data_filter = qfit_pos),
  "Colorectal cancer (inc sec care) if qFIT null" = first_table_row(any_crc_in_window, data_filter = qfit_three_cat == "null"),
  "FT referral" = ft_referral,
  # "Colonoscopy SUS" = !is.na(colonoscopy_sus),
  "Colonoscopy SUS if qFIT < 10" = first_table_row(!is.na(colonoscopy_sus), data_filter = !qfit_pos),
  "Colonoscopy SUS if qFIT >= 10" = first_table_row(!is.na(colonoscopy_sus), data_filter = qfit_pos),
  "Colonoscopy SUS if null qFIT" = first_table_row(!is.na(colonoscopy_sus), data_filter = qfit_three_cat == "null"),
  "Colonoscopy SUS if no qFIT" = first_table_row(!is.na(colonoscopy_sus), data_filter = is.na(qfit_date)),
  # "Flexi sig SUS" = !is.na(flexi_sig_sus),
  "Flexi sig SUS if qFIT < 10" = first_table_row(!is.na(flexi_sig_sus), data_filter = !qfit_pos),
  "Flexi sig SUS if qFIT >= 10" = first_table_row(!is.na(flexi_sig_sus), data_filter = qfit_pos),
  "Flexi sig SUS if null qFIT" = first_table_row(!is.na(flexi_sig_sus), data_filter = qfit_three_cat == "null"),
  "Flexi sig SUS if no qFIT" = first_table_row(!is.na(flexi_sig_sus), data_filter = is.na(qfit_date)),
  # "Colonoscopy primary care if qFIT < 10" = first_table_row(!is.na(colonoscopy_primary_care), data_filter = !qfit_pos),
  # "Colonoscopy primary care if qFIT >= 10" = first_table_row(!is.na(colonoscopy_primary_care), data_filter = qfit_pos),
  # "Colonoscopy primary care if null qFIT" = first_table_row(!is.na(colonoscopy_primary_care), data_filter = qfit_three_cat == "null"),
  # "Colonoscopy primary care if no qFIT" = first_table_row(!is.na(colonoscopy_primary_care), data_filter = is.na(qfit_date)),
  # "Flexi sig primary care if qFIT < 10" = first_table_row(!is.na(flexi_sig_primary_care), data_filter = !qfit_pos),
  # "Flexi sig primary care if qFIT >= 10" = first_table_row(!is.na(flexi_sig_primary_care), data_filter = qfit_pos),
  # "Flexi sig primary care if null qFIT" = first_table_row(!is.na(flexi_sig_primary_care), data_filter = qfit_three_Cat == "null"),
  # "Flexi sig primary care if no qFIT" = first_table_row(!is.na(flexi_sig_primary_care), data_filter = is.na(qfit_date)),
  # "Colonoscopy either" = !is.na(earliest_colonoscopy),
  # "Colonoscopy either if qFIT < 10" = first_table_row(!is.na(earliest_colonoscopy), data_filter = !qfit_pos),
  # "Colonoscopy either if qFIT >= 10" = first_table_row(!is.na(earliest_colonoscopy), data_filter = qfit_pos),
  # "Colonoscopy either if no qFIT" = first_table_row(!is.na(earliest_colonoscopy), data_filter = is.na(qfit_pos)),
  # "Flexi sig either" = !is.na(earliest_flexi_sig),
  # "Flexi sig either if qFIT < 10" = first_table_row(!is.na(earliest_flexi_sig), data_filter = !qfit_pos),
  # "Flexi sig either if qFIT >= 10" = first_table_row(!is.na(earliest_flexi_sig), data_filter = qfit_pos),
  # "Flexi sig either if no qFIT" = first_table_row(!is.na(earliest_flexi_sig), data_filter = is.na(qfit_pos)),
  # "Colonoscopy or flexi sig either" = !is.na(earliest_colonoscopy) | !is.na(earliest_flexi_sig),
  # "Colonoscopy either if qFIT < 10" = first_table_row(!is.na(earliest_colonoscopy) | !is.na(earliest_flexi_sig), data_filter = !qfit_pos),
  # "Colonoscopy either if qFIT >= 10" = first_table_row(!is.na(earliest_colonoscopy) | !is.na(earliest_flexi_sig), data_filter = qfit_pos),
  # "Colonoscopy either if no qFIT" = first_table_row(!is.na(earliest_colonoscopy) | !is.na(earliest_flexi_sig), data_filter = is.na(qfit_pos)),
  "Referral qFIT < 10" = first_table_row(ft_referral, data_filter = !qfit_pos),
  "Referral qFIT >= 10" = first_table_row(ft_referral, data_filter = qfit_pos),
  "Referral null qFIT" = first_table_row(ft_referral, data_filter = qfit_three_cat == "null"),
  "Referral no qFIT" = first_table_row(ft_referral, data_filter = is.na(qfit_pos)),
  "qFIT refl timing" = first_table_row((qfit_date %--% ft_referral_date) / ddays(1), data_filter = !is.na(qfit_date)),
  "qFIT cscopy timing" = first_table_row((qfit_date %--% earliest_colonoscopy) / ddays(1), data_filter = !is.na(qfit_date)),
  "qFIT flexi timing" = first_table_row((qfit_date %--% earliest_flexi_sig) / ddays(1), data_filter = !is.na(qfit_date)),
  "qFIT diagnosis timing" = first_table_row((qfit_date %--% crc_diagnosis_date) / ddays(1), data_filter = !is.na(qfit_date))
)
first_table(
  primary_data,
  .options = ft_options,
  .column_variable = region,
  !!!qfit_overall_rows
) %>% 
  print_ft()
```

## Main analysis, restricted to >=50 years old

```{r}
first_table(
  primary_data_over_50,
  .options = ft_options,
  .column_variable = region,
  !!!qfit_rows
) %>% 
  print_ft()
```

# Cumulative proportion of population having had a qFIT over time

Cumulative numbers of qFITs rounded to nearest 5 to avoid statistical disclosure control issues

```{r}
make_cum_prop_data <- function(data) {
  cum_props <- data %>% 
    filter(!is.na(qfit_date)) %>% 
    count(region, month = floor_date(qfit_date, "month")) %>%
    arrange(region, month) %>%
    group_by(region) %>%
    mutate(cum_num = cumsum(n)) %>%
    inner_join(
      data %>% count(region, name = "denom"),
      by = "region"
    ) %>%
    mutate(
      cum_num = {round(cum_num / 5, 0) * 5} %>% 
        {if_else(. == 5, 0, .)},
      cum_prop = cum_num / denom,
      across(c(cum_num, cum_prop), list(red = ~if_else(. == 0, "[REDACTED]", as.character(.))))
    )
  list(
    cum_props = cum_props,
    plot = cum_props %>% 
      ggplot(aes(month, cum_prop, group = region, colour = region)) +
      geom_line() +
      geom_point() +
      xlab("Month") +
      ylab("Cumulative proportion")
  )
}

cum_prop_data_over_50 <- make_cum_prop_data(input %>% filter(age >= 50))
cum_prop_data_all_ages <- make_cum_prop_data(input)

cum_prop_data_over_50$cum_props %>%
  select(-c(n, cum_num, cum_prop)) %>%
  rename_all(str_replace, "_red$", "") %>% 
  write_csv("../output/cum_prop_data.csv")

cum_prop_data_all_ages$cum_props %>%
  select(-c(n, cum_num, cum_prop)) %>%
  rename_all(str_replace, "_red$", "") %>% 
  write_csv("../output/cum_prop_data_all_ages.csv")
```

## Over 50's

```{r}
cum_prop_data_over_50$plot
ggsave("../output/cum_prop_graph.png", width = 12, height = 7)
```

## Over 50's

```{r}
cum_prop_data_all_ages$plot
ggsave("../output/cum_prop_graph_all_ages.png", width = 12, height = 7)
```

# Comparison of primary and secondary care CRC diagnosis (based on secondary admissions)

```{r}
primary_data_over_50 %>%
  count(
    region,
    primary_care_ever = earliest_crc_diagnosis_ever,
    primary_care_now = crc_diagnosis,
    secondary_care_ever = !is.na(earliest_crc_admission_ever_date),
    secondary_care_now = !is.na(crc_admission_date)
  ) %>%
  mutate(n = if_else(between(n, 1, 5), NA_integer_, n)) %>% 
  knitr::kable()
```

# Numbers by region and time period

```{r}
input %>% 
  group_by(region) %>% 
  summarise(
    n = n(),
    across(matches("qfit_p\\d+_date_present"), list(n = sum, prop = mean))
  ) %>% 
  rename_all(str_replace, "_date", "")
```