---
title: "qFIT OpenSafely analysis"
author: "Nick Kennedy"
date: "12/07/2021"
output:
  html_document:
      toc: true
      theme: cerulean
      df_print: kable
      dev: png
      fig_width: 12
      fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(rlang)
source("sro_gastro_functions.R")
walk(list.files("firsttable/R", full.names = TRUE), source)
flip_first_table <- function(x) {
  x %>% as.data.frame() %>%
    select(-Level) %>%
    column_to_rownames("Variable") %>%
    t() %>%
    as.data.frame()
}
```

```{r load data}
stp_codes <- read_csv("stp_codes.csv")
input <- read_csv("../output/input.csv", col_types = cols(.default = col_character())) %>% 
  mutate(
    across(matches("^(qfit|ft.*|crc.*)_date$"), lubridate::ymd),
    across(c(ft_referral, ft_clinic, crc_diagnosis), ~as.logical(as.integer(.x))),
    across(c(qfit, age, fobt_num), as.numeric),
    qfit = if_else(is.na(qfit_date), NA_real_, qfit),
    qfit_pos = qfit > 10,
  ) %>% 
  left_join(stp_codes, by = "stp")
```

# Top level summary by STP

```{r}
ft_options <- first_table_options(include_denom = TRUE, percent_first = TRUE, include_p = FALSE)
qfit_rows <- quos(
  "qFIT done" = !is.na(qfit_date),
  "qFIT > 0" = qfit > 0,
  "qFIT > 10" = qfit_pos,
  "Age" = age,
  "Age if qFIT done" = first_table_row(age, data_filter = !is.na(qfit_date)),
  "Age if qFIT neg" = first_table_row(age, data_filter = !qfit_pos),
  "Age if qFIT pos" = first_table_row(age, data_filter = qfit_pos),
  "Colorectal cancer" = crc_diagnosis,
  "Colorectal cancer if qFIT <= 10" = first_table_row(crc_diagnosis, data_filter = !qfit_pos),
  "Colorectal cancer if qFIT > 10" = first_table_row(crc_diagnosis, data_filter = qfit_pos),
  "FT referral" = ft_referral,
  "Colonoscopy" = !is.na(colonoscopy_sus),
  "Flexi sig" = !is.na(flexi_sig_sus),
  "Referral qFIT <= 10" = first_table_row(ft_referral, data_filter = !qfit_pos),
  "Referral qFIT > 10" = first_table_row(ft_referral, data_filter = qfit_pos),
  "Referral no qFIT" = first_table_row(ft_referral, data_filter = is.na(qfit_pos)),
  "qFIT refl timing" = (qfit_date %--% ft_referral_date) / ddays(1),
  "qFIT cscopy timing" = (qfit_date %--% colonoscopy_sus) / ddays(1),
  "qFIT flexi timing" = (qfit_date %--% flexi_sig_sus) / ddays(1),
  "qFIT diagnosis timing" = (qfit_date %--% crc_diagnosis_date) / ddays(1)
)
first_table(
  input,
  .options = ft_options,
  .column_variable = stp_name,
  !!!qfit_rows
) %>% 
  flip_first_table()
```

## Restricted to <50 years old

```{r}
first_table(
  input %>% filter(age < 50),
  .options = ft_options,
  .column_variable = stp_name,
  !!!qfit_rows
) %>% 
  flip_first_table()
```

## Restricted to >=50 years old

```{r}
first_table(
  input %>% filter(age >= 50),
  .options = ft_options,
  .column_variable = stp_name,
  !!!qfit_rows
) %>% 
  flip_first_table()
```

## Further data on qFIT by STP

```{r}
input %>%
  group_by(stp_name) %>% 
  summarise(
    n_people = n(),
    age_overall = med_iqr(age),
    female = sum(sex == "F"),
    had_qfit = sum(!is.na(qfit_date)),
    qfit_result = med_iqr(qfit[!is.na(qfit_date)]),
    qfit_gt_10 = sum(qfit > 10, na.rm = TRUE),
    had_ft_colorectal_referral = sum(ft_referral, na.rm = TRUE),
    had_ft_colorectal_clinic = sum(ft_clinic, na.rm = TRUE),
    had_colonoscopy = sum(!is.na(colonoscopy_sus)),
    had_flexi_sig = sum(!is.na(flexi_sig_sus)),
    had_crc_diagnosis = sum(crc_diagnosis, na.rm = TRUE),
    referral_if_qfit_gt10 = sum(!is.na(ft_referral_date[qfit > 10 & !is.na(qfit_date)])),
    referral_if_qfit_le10 = sum(!is.na(ft_referral_date[qfit <= 10 & !is.na(qfit_date)])),
    referral_if_qfit_not_done = sum(!is.na(ft_referral_date[is.na(qfit_date)])),
    qfit_referral_timing = med_iqr((qfit_date %--% ft_referral_date) / ddays(1), na.rm = TRUE),
    qfit_clinic_timing = med_iqr((qfit_date %--% ft_clinic_date) / ddays(1), na.rm = TRUE),
    qfit_colonoscopy_timing = med_iqr((qfit_date %--% colonoscopy_sus) / ddays(1), na.rm = TRUE),
    qfit_flexi_sig_timing = med_iqr((qfit_date %--% flexi_sig_sus) / ddays(1), na.rm = TRUE),
    qfit_diagnosis_timing = med_iqr((qfit_date %--% crc_diagnosis_date) / ddays(1), na.rm = TRUE),
    qfit_age = med_iqr(age[!is.na(qfit_date)]),
    qfit_female = sum(sex[!is.na(qfit_date)] == "F"),
    top_four_fit_values = table(round(qfit[!is.na(qfit_date)], 0)) %>%
      sort(decreasing = TRUE) %>%
      head(4) %>% 
      sprintf("%s:%d", names(.), .) %>%
      paste(collapse = "; ")
  )
```